A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-10-24 paper/elsarticle/elsarticle-template.tex
paper

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection, in units of pixels, of $\mathbf{X}_{i,j}$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. More generally, the projection $\mathbf{x}' \in \mathbb{R}^2$ of $\mathbf{X}$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
\begin{equation}
(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
\end{equation}
where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses the point~$\mathbf{X}$ in the coordinate frame of the camera. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

@@

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location. (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector. This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units. Note that pixels shown in black will not be projected to the scene.}
    \label{fig:warp}
\end{figure}

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.

>>>>

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection of the point~$\mathbf{X}_{i,j} \in \mathbb{R}^3$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$. The function~$d_{\hat{\omega}}$, in turn, is a function that first (i) corrects $\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'$ for lens distortions in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}, and then (ii) computes a distance in pixels with respect to two resulting corrected positions. More generally, the projection~$\mathbf{x}' \in \mathbb{R}^2$ of a point~$\mathbf{X} \in \mathbb{R}^3$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
\begin{equation}
(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
\end{equation}
where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses $\mathbf{X}$ in the coordinate frame of the camera. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

@@

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location. (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector. This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired metric dimensions (i.e., desired scale) in the scene plane. Note that pixels shown in black will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.
